Writing function in R
========================================================

This document is on writing function in R. It contains elements:
* defining arguments
* passing arguments to function
* checking arguments
* validation and stoping
* errors and warnings
* useful functions (parse, eval,substitute, deparse, args, match.call,flush.console())

In this tutorial we would like to make function that test correlation and procudes heat map using ggplot2 package. For this we need to install following packages: ggplot2, reshape2. We would like use also non-standard correlation tests that were discussed in [this post](http://smarterpoland.pl/index.php/2014/02/korelacje/) @smarterpoland.pl.

* energy
* HHG
* minerva
* Hmisc

```{r instalingpackages,eval=F}
install.packages(c('rococo','energy','HHG','minerva','Hmisc'))
```

```{r loadingPackages,echo=FALSE,warning=F,error=FALSE}
library(energy)
library(HHG)
library(minerva)
library(Hmisc)
```

1. Defining arguments
---------------

For simple correlation we could use cor function (to test significance cor.test). First we need to load data.

```{r loadData}
load("~/Dokumenty/Projekty/Projekty_R/UNIVERSITY/Course Materials/REAL ESTATE/DataSets/RData/realEstatesKWW.RData")
### subset only to Wroclaw
Wroc<-realEstatesKWW[realEstatesKWW$city=='Wroclaw',]
```

Check correlation using cor function

```{r}
names(Wroc)
cor(Wroc[,c('nRooms','surface','apprPrice','appPriceM2')],method='spearman')
## this code also works cor(Wroc[,c('nRooms','surface','apprPrice','appPriceM2')],method='s')
cor(Wroc[,c('nRooms','surface','apprPrice','appPriceM2')],method='pearson')
```










