\documentclass[12pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[polish,english]{babel}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\author{Maciej Beręsewicz}
\title{Real Estate Market -- a glossary}
\begin{document}
\maketitle
\begin{description}
\item[Active sewage network] (\textit{Sieć kanalizacyjna czynna}) A system of covered canals (underground) discharging sewage from buildings and other objects to collectors or wastewater treatment facilities.
\item[All-navigable sewage network] (\textit{Sieć kanalizacyjna ogólnospławna}) Sewers designed for combined draining of all types of wastewater, i.e. household, industrial, rainfall or thaw.
\item[Apartment heating.] (\textit{Ogrzewający mieszkanie}) Households consuming gas for heating purposes.
\item[Arrears in payments for dwellings] (\textit{Zaległości w opłatach za mieszkanie}) The amount not paid both by tenants and owners-occupiers of dwellings, due to charges for the used dwellings, i.e. rent, water-rate, charge on waste collection, lift, etc.
\item[Arrears in repayment of /housing loan] (\textit{Zaległości w spłacie kredytu mieszkaniowego}) The amount of installments in arrears (including interest), which debtors should pay into the housing co-operative bank account.
\item[Boiler house] (\textit{Kotłownia}) A boiler-house is a building or a room with boilers and devices used for production of thermal energy for heating or both heating and supplying warm water.
\item[Building equipped with central heating] (\textit{Budynek wyposażony w centralne ogrzewanie}) A building in which all or some of dwellings are connected to an installation delivering heat from a central source of its production as well as a boiler house in an individual building.
\item[Building equipped with gas from gas supply network] (\textit{Budynek wyposażony w gaz z sieci}) A building in which all or some of dwellings are equipped with gas supply system.
\item[Building equipped with sewerage network] (\textit{Budynek wyposażony w kanalizację}) A building inside of which - in dwellings or in the corridor - there is installation (to which devices like: sink, flushed toilet, washbasin, bathtub are connected), enabling drainage of waste and waste water to collectors.
\item[Building fitted with water supply network] (\textit{Budynek wyposażony w wodociąg}) A building in which all or some dwellings are equipped with a tap with running water or in case when in the building there are not dwellings with running water but it is situated outside the dwelling but in the building, e.g in the passage.
\item[Building owned by company] (\textit{Budynek stanowiący własność zakładu pracy}) Building which is in whole (i.e. no dwelling in it constitutes separate ownership of other entities) owned by a company - organizational unit holding ownership title to this building, certified by a record in land and mortgage register, and in case there is no register, in other document certifying the ownership.
\item[Building owned by condominium] (\textit{Budynek stanowiący współwłasność}) Building having several legal owners found in land and mortgage register.
\item[Building owned by gmina] (\textit{Budynek stanowiący własność gminy}) Building which in whole (i.e. no dwelling in them constitutes separate ownership of other entities: - are owned by a gmina, remaining under management of dwelling or municipal entities established specifically to manage gmina's dwelling stock, as well as buildings under direct management of gmina's office, - were handed over to a gmina, but remaining at the disposal of units of public use such as: health care centers, units of education system, institutions of science and culture.
\item[Building owned by housing co-operative] (\textit{Budynek stanowiący własność spółdzielni mieszkaniowej}) Building fully owned by a housing co-operative and in which there is no separate ownership of premises by another entity.
\item[Building owned by natural person] (\textit{Budynek stanowiący własność osoby fizycznej}) Building constituting the unlimited property of a natural person - a single person or both spouses (i.e., not containing any premises that would constitute the separate property of another entity) and registered in the land and mortgage registry or, in case of the lack of such a registry, in another document confirming ownership rights, all regardless of whether or not the owner lives in such a building or has rented it or given it to another person (persons) free of charge for use or the building has been given approval for use but has not yet been inhabited.
\item[Building owned by public building societies] (\textit{Budynek stanowiący własność towarzystwa budownictwa społecznego}) Building owned in whole by a legal entity (i.e. no dwelling in them constitutes separate ownership of other entity), which in their names have the words: "public building society" or Polish acronym TBS, irrespective of participating in costs of constructing a building of other entity (a gmina, a housing co-cooperative, a company) in return for receiving the dwellings (on rental terms) for third parties indicated by these entities.
\item[Building owned by the State Treasury] (\textit{Budynek stanowiący własność Skarbu Państwa}) Building which in whole (i.e. no dwelling in them constitutes separate ownership of other entity) remains under management of entities administrated property of State Treasury.
\item[Buildings owned by other entity] (\textit{Budynek stanowiący własność innego podmiotu}) Building in whole (i.e. no dwelling in them constitutes separate ownership of other entities) owned byentities not covered by terms: "Building owned by natural person", "Building owned by housing co-operative", "Building owned by gmina", "Building owned by the State Treasury", "Building owned by company", "Building owned by public building societies".
\item[Business premises] (\textit{Lokal użytkowy}) Premises used only for economic activity on rental terms and in the buildings owned by condominiums also premises constituting a separate ownership (retail outlets, service premises, production premises, studios, premises rented as office spaces, premises used for cultural and social activity etc.), which are a source of profit for the entity that rents them.
\item[Capacity production of water supply facilities of twenty-four hours] (\textit{Dobowa zdolność produkcyjna urządzeń wodociągowych}) The amount of water which can be produced during 24 hours, considering the capabilities of production facilities at the end of year, without disturbing the equilibrium of water intakes and with preservation of technical conditions and requirements.
\item[Collective accommodation building] (\textit{Budynek zbiorowego zakwaterowania}) A building in which more than half of the surface area is occupied by multi-person premises (e.g., boarding house, student dormitory, children's home, sanatorium) and in which at least one dwelling is located.
\item[Collective accommodation places] (\textit{Obiekt zbiorowego zakwaterowania}) A group of premises (rooms and other ancillary premises) located in one or more buildings and occupied by one separate entity rendering the following types of services: hotel, care and education, healthcare or other in which the services rendered require that a larger number of people live in such objects.
\item[Condominium] (\textit{Wspólnota mieszkaniowa}) A multi-dwelling building (or a few buildings), in which a part or all dwellings constitute separate properties of natural persons, certified by a record in land and mortgage register.
\item[Consumption of gas in households] (\textit{Zużycie gazu w gospodarstwach domowych}) Gas delivered to households.
\item[Cost of provided services] (\textit{Koszty świadczonych usług}) Costs connected with providing services for residential and business premises i.e. supply of thermal energy (central heating, hot water), cold water, sewage collection, liquid and solid waste collection, and maintenance of lifts (provided they are not treated as maintenance and renovation costs).
\item[Costs of dwelling stock maintenance] (\textit{Koszty utrzymania zasobów lokalowych}) Costs borne for dwelling stock maintenance and business premises i.e. operating costs and costs of provided services.
\item[Decrements of dwelling stock] (\textit{Ubytki zasobów mieszkaniowych}) Physical decrements specified based on reports on demolition of buildings, decisions on demolition enforcement, issued permits to change the purpose of a residential building or a dwelling into purposes other than residential, decisions on enforcement of demolition of buildings not used, damaged or not finished, unfit for renovation, reconstruction or finishing.
\item[Distribution thermic-line] (\textit{Sieć cieplna rozdzielcza}) System of distribution pipes supplying heating factor to household connections.
\item[Dwelling] (\textit{Mieszkanie}) A premise for a permanent residence of persons - built or remodelled for residential purposes structurally separated by fixed walls within a building, into which a separate access leads from a staircase, a passage, a common hall or directly from the street, courtyard or garden, comprising one or several rooms and auxiliary spaces.
\item[Dwelling equiped with toilet] (\textit{Mieszkanie wyposażone w ustęp}) A dwelling fitted with a sanitary system within its premises, connected to the water supply system, and discharge waste water to the sewage system, or to the local appliances (septic tanks).
\item[Dwelling fitted with a bathroom] (\textit{Mieszkanie wyposażone w łazienkę}) A dwelling, in which there is a room, with a bathtub or a shower (or both of them), as well as a system channeling waste water to the sewage system, or to the local appliances (septic tanks).
\item[Dwelling fitted with central heating] (\textit{Mieszkanie wyposażone w centralne ogrzewanie}) A dwelling fitted with a system within the premises, which supplies heat from a central heating source, i.e. heat and power stations, thermal power station, local boiler house within the premises of the housing estate, central heating furnace installed in own boiler house or in any other room. Electric floor heating is also regarded as central heating.
\item[Dwelling fitted with gas] (\textit{Mieszkanie wyposażone w gaz}) A dwelling fitted with a system within the premises, which (along with recipient installations in the dwelling) supplies gas to active connections.
\item[Dwelling fitted with running hot water] (\textit{Mieszkanie wyposażone w ciepłą wodę}) A dwelling to which hot water is supplied from a combined heat and power plant, heat plant, a housing estate boiler-room or a local boiler-room in a multi-dwelling building. It is also a dwelling in which water is heated using internal appliances, such as: a gas water-heater, a water heater, a boiler - connected to the water supply system.
\item[Dwelling fitted with running hot water supplied from heat and power plant, local boiler house] (\textit{Mieszkanie wyposażone w ciepłą wodę dostarczaną z elektrociepłowni, ciepłowni  lub kotłowni osiedlowej}) A dwelling where hot water is supplied form a combined hesat and power plant, heat plant, a housing estate boiler-room and distributed by means of the installation (to the intake facilities located in the dwelling).
\item[Dwelling fitted with sewage system] (\textit{Mieszkanie wyposażone w kanalizację}) A dwelling fitted with sewage installation (in which there is a sanitary system) which discharge sewage into the sewage system or local sewage system (septic tanks).
\item[Dwelling fitted with sewage system] (\textit{Mieszkanie wyposażone w kanalizację z sieci}) A dwelling fitted with sewage installation (in which there is a sanitary system) which discharge sewage into the sewage system.
\item[Dwelling fitted with water supply system] (\textit{Mieszkanie wyposażone w wodociąg}) A dwelling fitted with a tap with running water within its premises. 'Supply of piped water' is understood as a system (including the recipient installations in the dwelling), which supplies water from the waterworks (by means of active connections) from the street pipeline to local systems (own water intake).
\item[Dwelling fitted with water supply system] (\textit{Mieszkanie wyposażone w wodociąg z sieci}) A dwelling fitted with a tap with running water within its premises. 'Supply of piped water' is understood as a system (including the recipient installations in the dwelling), which supplies water from the waterworks.
\item[Dwelling owned by the State Treasury] (\textit{Mieszkanie stanowiące własność Skarbu Państwa}) Dwelling which remains under management of entities administrated property of State Treasury. 
\item[Dwelling stock] (\textit{Zasoby mieszkaniowe}) Total number of inhabited and uninhabited dwellings in residential and non-residential buildings.
\item[Dwelling temporary inhabited] (\textit{Mieszkanie zamieszkane czasowo}) A dwelling temporary inhabited by persons registered  during Census Population, but none of them was not deemed as a stationary inhabitant in this area.
\item[Dwellings fitted with (sanitary and technical) installations] (\textit{Mieszkanie wyposażone w instalacje sanitarno-techniczne}) Dwellings with at least one of the following sanitary and technical appliances: supply of piped water, a flushable toilet, hot water, a bathroom, central heating or gas supply from the gas pipeline.
\item[Dwellings owned by a gmina] (\textit{Mieszkania stanowiące własność gminy}) Dwellings- in buildings that are in whole owned by a gmina and individual dwellings owned by a gmina, but located in buildings constituting a shared real property, i.e. dwellings which serve meeting the needs of all inhabitants of a gmina; remaining in administration of a municipal organizational unit established in order to administrate municipal dwelling stock (e.g. Housing Management Enterprise/Institution, Municipal Housing Administration) or remaining in direct administration of a Community Office,- owned by the State Treasury but administrated by a municipality,- handed over to a gmina, but remaining at the disposal of public use units, such as: health care institutions/centres (ZOZ), units of the education system, institutions of science and culture etc. serving mainly the housing needs of employees of such units. In statistical surveys, dwellings owned by gminas also include dwellings owned by powiats (local powiat self-government communities), but not being the property of any gmina, but remaining in administration of that municipality or a municipal organizational unit established in order to administrate dwelling stock.
\item[Dwellings owned by companies] (\textit{Mieszkania stanowiące własność zakładów pracy}) Dwelling in whole owned by a company.
\item[Dwellings owned by housing co-operatives] (\textit{Mieszkania stanowiące własność spółdzielni mieszkaniowych}) Dwellings in buildings which are the property or joint property of housing co-operatives, excluding dwellings for which, on the grounds of the Act of December 15, 2000 on housing co-operatives, a separate ownership title was established, for the benefit of one or more natural persons.
\item[Dwellings owned by natural persons] (\textit{Mieszkania stanowiące własność osób fizycznych}) Dwellings, the ownership title to which belongs to a natural person (one or more, e.g. spouses), while this person: - can be the owner of the whole real estate in which the dwelling is situated; for example an individual single family house, - can have a share in a shared real estate, as the right related to the separate ownership title to a dwelling, e.g. located in a building included in a condominium.
\item[Dwellings owned by other entities] (\textit{Mieszkania stanowiące własność innych podmiotów}) Dwellings constituting the property of:1) private organizations involved in the building or buying of dwellings for profit: for sale or for rental;2) trade unions, associations, foundations, political parties, professional and economic self-governments;3) the Roman-Catholic Church and other churches and religious associations, including catholic universities and church institutes,4) other entities, not covered by terms: "Dwellings owned by natural persons" "Dwellings owned by housing co-operatives" "Dwellings owned by companies" "Dwellings owned by a gmina" "Dwellings owned by the State Treasury" "Dwellings owned by public building societies"
\item[Dwellings owned by public building societies] (\textit{Mieszkania stanowiące własność towarzystw budownictwa społecznego}) Dwellings in buildings being the property of legal entities having in their names "public building society" or the Polish abbreviation "TBS" irrespective of participation in costs of constructing the building by another entity (a gmina, a housing co-operative, a company) in return for receiving the dwellings (on rental terms) for third parties indicated by such an entity.
\item[Eviction] (\textit{Eksmisja}) All actual and legislative activity aimed at removing out a person from occupied dwelling or property (to vacate from persons end things).
\item[Floor space of dwellings fitted with central heating] (\textit{Powierzchnia mieszkania ogrzewanego centralnie}) Floor space of dwellings located in residential and non-residential buildings connected to heating system or with local boiler-rooms.
\item[Gas recipient.] (\textit{Odbiorca gazu}) Anyone who receives or draws gas fuels under a contract with the energy company.
\item[Gas sale] (\textit{Sprzedaż gazu}) Gas sold by energy companies (sales = gas consumption).
\item[Gas supply connection] (\textit{Przyłącze do sieci gazowej}) Segment of network from the supplying gas pipeline to separated armature, used for connecting to gas supply network equipments or the gas installation join entity.
\item[Gas supply distribution network] (\textit{Sieć rozdzielcza gazowa}) The gas supply network used for transmission and distribution of gas fuels with pressure not higher than 0.5MPa.
\item[Gas transmission network] (\textit{Sieć przesyłowa gazowa}) Gas network for transmission and distribution of gas at a pressure above 0.5 MPa.
\item[Gas user] (\textit{Użytkownik gazu}) Anyone who uses gas with or without an agreement with the energy company.
\item[Gas users under the contract.] (\textit{Korzystający z gazu na podstawie umowy.}) Households consuming gas on the basis of an individual contract with energy company.
\item[Gas users without a contract.] (\textit{Korzystający z gazu bez umowy.}) Households consuming gas without a contract with energy company (collective agreements between housing association or a community and energy company).
\item[Heating an apartment without a contract.] (\textit{Ogrzewający mieszkanie bez umowy.}) Households consuming gas without a contract with energy company also for heating purposes.
\item[Heating apartment under the contract.] (\textit{Ogrzewający mieszkanie na podstawie umowy.}) Households consuming gas (on the basis of an individual contract with energy company) also for heating purposes.
\item[Heating connection] (\textit{Przyłącze ciepłownicze}) Section of heating network supplying heat only to one district heating or a section of external installation over district heating substation or heat source, joining these installations with receiving installations in buildings.
\item[Holiday (recreational) home] (\textit{Budynek rekreacyjny}) A building which construction (pursuant to the building law) require permits as for residential buildings but in which the main purpose of stays is temporary, e.g., during vacation time, days off from work, as well as a building that was previously a building for permanent living but has been adapted for temporary living.
\item[Housing allowance] (\textit{Dodatek mieszkaniowy}) Housing allowance is a common and temporary form of financial aid. The amount of housing allowance is the difference between the housing expenses incurred by the standard usable floor space of an occupied dwelling unit and that part of the expenses incurred by the person to whom the benefit is granted.
\item[Independent residential apartment] (\textit{Samodzielny lokal mieszkalny}) A residential apartment, in relation to which the starost has determined, in the form of a statement, the compliance with the requirements mentioned in article 2 section 2 of the act of June 24, 1994 on the ownership of premises.
\item[Infrastructure] (\textit{Infrastruktura}) The system of public works in a country, state or region, including roads, utility lines and public buildings. 
\item[Inhabited dwelling] (\textit{Mieszkanie zamieszkane}) A dwelling in which at least one person resides, regardless of whether such a person is registered for a temporary or permanent residence in this dwelling or stays in it without being registered for a temporary or permanent residence.
\item[Inhabited space which is not a dwelling] (\textit{Zamieszkane pomieszczenie nie będące mieszkaniem}) Inhabited space which is not a dwelling contains premises not adapted to residential purposes and inhabited temporarily,  a mobile unit or other units which are the one place of living for person during Census Population.
\item[Kitchen] (\textit{Kuchnia}) A premise within a dwelling (with or without a window) that is separated by fixed walls, designated towards and appropriately equipped to prepare meals and mainly used for this purpose regardless of whether such premises are also used to consume meals, to sleep in or for other household purposes.
\item[Kitchen space] (\textit{Pomieszczenie kuchenne}) Place for the preparation of meals irrespective of whether it is also used for eating, sleeping or living.
\item[Losses from water supply network] (\textit{Straty wody w sieci wodociągowej}) The amount of water which escaped as a result of leakage or breakdown.
\item[Loyal little for occupation of dwelling by household] (\textit{Tytuł prawny do zajmowania mieszkania przez gospodarstwo domowe}) A legal claim to occupation of dwelling is set for every group of persons that constitutes a household. In terms of rights to the occupied dwelling households can be classified as those occupying a dwelling by virtue of: ownership of a dwelling or a house occupancy rights to a specific suite within the housing co-operative lease sublease being related to an owner or a main occupant of a dwelling and other.
\item[Major overhauls] (\textit{Remont kapitalny}) Renovation work aiming at restoring buildings to the original or almost original technical and utility value.
\item[Non-residential building] (\textit{Budynek niemieszkalny}) A building in which more than half of the surface area is occupied by non-residential premises (e.g., school, offices, shop, warehouse, medical centre) and in which at least one dwelling is located.
\item[Open gas connections] (\textit{Czynne przyłącza gazowe}) Connection filled with gas (from the pipeline supply to the main tap including protection ) to a residential building in which at least one gas recipient of the gas has an agreement for the purchase  sale of gas.
\item[Open gas connections to residential buildings] (\textit{Czynne przyłącza gazowe do budynków mieszkalnych.}) Connection filled with gas  (from the pipeline supply to the main tap including protection ) to a residential building in which at least one gas recipient of the gas has an agreement for the purchase  sale of gas.
\item[Operating costs] (\textit{Koszty eksploatacji}) Costs connected with dwelling stock maintenance and business premises.
\item[Permanently inhabited dwelling (with actual inhabitants)] (\textit{Mieszkanie zamieszkane stale (z faktycznymi mieszkańcami)}) A dwelling inhabited by at least one person registered for a permanent residence in this dwelling, even if they are absent for a long time (even a few years), e.g. is abroad or in other area in Poland because of an education, a job, a military service, being in a social care home or in a prison etc.
\item[Premises] (\textit{Lokal}) The room or set of rooms separated with the durable walls within the building dedicated to the permanent stay of people, which together with the auxiliary rooms serve the purpose of fulfilling their housing needs or which are used according to their dedication for the purposes other than residential purposes.
\item[Real estate turnover] (\textit{Obrót nieruchomościami}) A transfer of property of premises, buildings, or land from a seller to a purchaser in form of a market transaction or non-market transaction. A real estate turnover transaction requires a notarial deed and to record in land and mortgage register.
\item[Register of real estate prices and values] (\textit{Rejestr Cen i Wartości Nieruchomości (RCiWN)}) The public register kept by a starost which include data on real estate prices given in notarial deeds and real estate values given by real estate appraisers in appraisal reports which extracts are provided to the register of land and buildings.
\item[Residential-livestock building or farm-residential building] (\textit{Budynek mieszkalno-inwentarski lub mieszkalno-gospodarski}) A building that consists of both a residential part including one or more dwellings as well as a farm part (consisting of a stable, cowshed, piggery, etc.) or ancillary services associated with the running of a farm regardless of what part of such a building is occupied by the residential part.
\item[Residential building] (\textit{Budynek mieszkalny}) A building designated towards residential purposes and fully occupied by residential premises or a building in which at least one half of the overall space constitutes residential space with the remaining part constituting other space (except for residential-animal buildings or residential-workshop buildings). 
\item[Room] (\textit{Izba}) A room is a space in a dwelling, separated from other spaces by permanent walls from the floor to the ceiling, with an area of at least 4 square metres, with direct day lighting, i.e. with a window or a French window in an external wall of the building; kitchens are also considered rooms as long as they fulfill the above criteria.
\item[Room (2)] (\textit{Pokój}) Premises located within a dwelling and fulfilling the definition of a room, built or adapted to housing purposes, e.g., sleeping, learning, relaxation with the exception of kitchen rooms.
\item[Sewage connection (1)] (\textit{Przyłącze kanalizacyjne}) The section of connection connecting the internal sewage system of a dwelling being the recipient of sewage network services, after the closest manhole looking from the building or, in case of the lack of such a manhole, to the border of the property.
\item[Sewage system] (\textit{Kanalizacja}) The complete sewage collection, treatment and disposal system: sewage networks, outlets of devices used to emit sludge into the waters or to the ground, sewage pretreatment facilities and sewage treatment plants, and sewage pumping stations.
\item[Sewage system installations] (\textit{Urządzenia kanalizacyjne}) Sewage networks, sewage discharge equipment used to discharge sewage into water or soil and preliminary sewage treatment and treatment equipment as well as sewage pumps.
\item[Social dwelling] (\textit{Mieszkanie socjalne}) A dwelling fit for residing in due to furnishings and the technical condition; the space of rooms of such a dwelling for a household dweller of the renter may not be smaller than 5 sq.m, and in case of single-person household must be at least 10 sq.m, whilst the standard of such a dwelling may be lowered.
\item[Street outlets] (\textit{Zdroje uliczne}) Publicly available facilities directly connected to street water mains, serving the community for drawing the water directly from such the mains.
\item[Substandard conditions in dwellings] (\textit{Substandardowe warunki mieszkaniowe}) Lowered housing conditions caused by the low technical condition of a building, an insufficient installation equipment and an overpopulation.
\item[Transmission (main) thermic-line] (\textit{Sieć cieplna przesyłowa (magistralna)}) System of pipes supplying heating agent to distribution pipes.
\item[Transmission (main) water supply network] (\textit{Sieć wodociągowa przesyłowa (magistralna)}) Conduits supplying water from a distant water intake to water supply distribution network.
\item[Uninhabited dwelling] (\textit{Mieszkanie niezamieszkane}) A dwelling in which at the moment of the statistical survey none person was registered (for a permanent or a temporary residence) nor any unregistered person stayed, even temporarily.
\item[Useful floor area of a dwelling] (\textit{Powierzchnia użytkowa mieszkania}) The total useful area of all premises in a dwelling or in a residential building in which there is only one dwelling, i.e. rooms, a kitchen, larders, anterooms, alcoves, halls, corridors, bathrooms, toilets, cased veranda, porch, dressing room and other premises for residential and utility needs of dwellers regardless of their designation and purpose of use (among others: ateliers, recreation facilities etc).
\item[Water collected from intakes] (\textit{Woda pobrana z ujęć}) The amount of water collected from an intake at the specific time, e.g. in a year, both raw water, requiring treatment (purification) and clean water not requiring purification.
\item[Water collected from surface intakes] (\textit{Woda pobrana z ujęć powierzchniowych}) Water collected from rivers, lakes and artificial water reservoirs.
\item[Water delivered for production purposes] (\textit{Woda dostarczona na cele produkcyjne}) Water delivered to industrial, construction, transportation, etc. enterprises (plants), i.e. production plants in all sectors of national economy, regardless of whether the water is used for technological purposes or for social and living needs of employees (in washbasins, bathrooms, dining-rooms, community rooms, office premises, etc. located in the plant).
\item[Water delivered to households] (\textit{Woda dostarczona gospodarstwom domowym}) Consumption of water from water supply system in households and group institutional households, against payment or free of charge, regardless of the amount of payment for such the water and the location of the household (town, village).
\item[Water supply connection] (\textit{Przyłącze wodociągowe}) The section of connection connecting the water supply network with the internal water network in the dwelling being the recipient of such services together with the valve on the main water meter.
\item[Water supply distribution network] (\textit{Sieć rozdzielcza wodociągowa}) Street conduits for distribution of water to recipients through connections leading to buildings and other objects.
\item[Water supply installations] (\textit{Urządzenia wodociągowe}) Ground level and underground water intakes, public wells, equipment used for the storage and treatment of water, waterworks networks, equipment regulating water pressure.
\item[Water supply system] (\textit{Wodociąg}) A set of water network devices serving collection of surface and underground waters, public wells, devices serving storage and treatment of water, water supply networks, water pressure control devices.
\item[Wholesale of water/consumption of water] (\textit{Sprzedaż hurtowa wody/zużycie wody}) Water sold (or transferred free of charge) to other entities, e.g. municipal water supply distribution stations, industrial plants etc) for further resale (in whole or in part) through an own network of these entities (other receivers, e.g. private households, plants), irrespective of rates paid by the receivers.
\item[Wholesale recipient.] (\textit{Odbiorca hutowy}) Entity (natural person or legal person holding a license for transmission, distribution and trade of gaseous fuels) purchasing gas in bulk for resale.
\item[Year of construction] (\textit{Rok wzniesienia budynku}) The year in which the construction was completed, i.e. the building was commissioned. If the building was commissioned in parts (individual staircases, extensions), the year of completing the last part of the construction is assumed.
\end{description}
\end{document}