---
title: "Data Quality"
author: "Maciej Beręsewicz"
date: "19 Apr 2015"
output: html_document
---

# Missing data analysis and imputation

Material contains basic information about missing data analysis and imputation. We start with installing the following packages

```{r install_packages,echo=TRUE,warning=FALSE, message=FALSE, error=FALSE}
library(mi) ### visualisation of missing data patterns (and multiple imputation)
library(Hmisc) ### impute and na.pattern function
library(VIM) ### visualisation of missing data, knn imputation
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(stringi)
library(lubridate)
```

Now read data from the Register of Prices and Values

```{r read_data_xlsx,echo=TRUE}
fname <- '~/Documents/Uczelnia/Rozne/GeoPoz - dane/ZG-OUG.41021.2.2.2015/Lokale_rynek_wtórny_2008.xls'
excel_sheets(fname)
dsin <- read_excel(fname,skip=2)
```

Check what types of dwellings we have in the data

```{r check_dwellings}
dsin <- dsin %>%
  rename(primary_function = `Funkcja podstawowa lokalu`)
dsin %>% count(primary_function)
```

We are interested in 'Mieszkalny' which is (residential) dwelling. 

```{r filter_dwellings}
dwellings <- dsin %>% 
  filter(stri_trim_both(primary_function) == 'Mieszkalny') %>%
  dplyr::select(`Nr transakcji`:`Cena za nieruchomość [PLN]`,
                `Adres i nr lokalu`,`Udział w prawie do lokalu`,
                `Liczba izb`:`Cena m2 lokalu [PLN]`) %>%
  rename(trans_id = `Nr transakcji`,
         trans_date  = `Data transakcji`,
         price_trans = `Cena transakcji [PLN]`,
         estate_id = `Nieruchomość`,
         estate_price = `Cena za nieruchomość [PLN]`,
         dwelling_address = `Adres i nr lokalu`,
         dwellings_rooms = `Liczba izb`,
         dwelling_floor_area = `Powierzchnia lokalu [m2]`,
         dwelling_floor_number = `Kondygnacja`,
         dwellong_assoc_facilities = `Liczba pomieszczeń przynależn.`,
         dwelling_assoc_facilities_flor_area = `Powierzchnia pom. przyn. [m2]`,
         dwelling_price = `Cena za lokal [PLN]`,
         dwelling_price_mw = `Cena m2 lokalu [PLN]`,
         participation_right = `Udział w prawie do lokalu`)
  
```


Change type of columns to see the missing values

```{r type_conversion}
dwellings <- dwellings %>%
  mutate_each(funs(as.numeric),
              price_trans,estate_price,dwellings_rooms:dwelling_price_mw) %>%
  mutate(trans_date = dmy(trans_date))
summary(dwellings)
```

Check missing data pattern

```{r missing_data_pattern}
dwellings %>% 
  dplyr::select(price_trans,estate_price,
                dwellings_rooms,dwelling_floor_area,dwelling_floor_number,
                dwelling_price,dwelling_price_mw) %>%
  na.pattern()
```

and visualise

```{r missing_data_pattern2}
dwellings %>% 
  dplyr::select(price_trans,estate_price,
                dwellings_rooms,dwelling_floor_area,dwelling_floor_number,
                dwelling_price,dwelling_price_mw) %>%
  aggr() %>%
  summary()
```

Check whether the floor missing vary in time

```{r floor_missing}
dwellings %>%
  mutate(ym = paste0(year(trans_date),'-',month(trans_date))) %>%
  group_by(ym) %>%
  summarise(rooms_miss = sum(is.na(dwelling_floor_number)),
            number_trans = n(),
            perc = rooms_miss/number_trans) %>%
  ungroup() %>%
  mutate(ym=paste0(ym,'-1'),
         ym = ymd(ym)) %>%
  arrange(ym)
```

Check whether the same flats have missing data

```{r}
dwellings %>% count(estate_id, sort=T)
dwellings %>%
  filter(estate_id == 'PO1P/00189541/6') 
```

Check duplicates (without trans id)

```{r duplicates}
dwellings[duplicated(dwellings[,-1]),]
```


